### 2022-07-27 paper/localpsf.tex
fixing another small typo
two more miniscule changes

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches. These approximations are labeled by are denoted by PSF (1), PSF (5), and PSF (25), respectively. The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$. We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point. Our results show that our preconditioners are also good preconditioners for $H$. The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$. Straightforward calculation shows that the discretized Hessian of the regularization term is given by
$
(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
$
where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques. We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.

@@

We presented an efficient matrix-free method for approximating operators with locally supported non-negative integral kernels. The method only requires access to the operator via application of the operator to a small number of vectors; fast access to entries of the operator's matrix representation is not required. The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these known impulse responses to approximate unknown entries of the operator's integral kernel. The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance. The ability to quickly approximate arbitrary integral kernel entries permits us to form a H-matrix approximation of the operator. Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization or inversion. The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure we presented that involves applying the operator to a small number of polynomial functions. Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the supports of the corresponding ellipsoids do not overlap. We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods. Although the method we presented in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.

>>>>

We use the framework described in this paper to generate Hessian preconditioners. We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches. These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively. The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$. We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point. Our results show that our preconditioners are also good preconditioners for $H$. The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$. Straightforward calculation shows that the discretized Hessian of the regularization term is given by
$
(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
$
where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix. These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques. We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.

@@

We presented an efficient matrix-free method for approximating operators with locally supported non-negative integral kernels. The method only requires access to the operator via application of the operator to a small number of vectors; fast access to entries of the operator's matrix representation is not required. The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these known impulse responses to approximate unknown entries of the operator's integral kernel. The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance. The ability to quickly approximate arbitrary integral kernel entries permits us to form a H-matrix approximation of the operator. Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion. The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure we presented that involves applying the operator to a small number of polynomial functions. Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the supports of the corresponding ellipsoids do not overlap. We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods. Although the method we presented in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.
